manuscript

### 2022-09-21 ms/manuscript.Rmd
update manuscript

---
title: "manuscript"
author: "Daniel Suh"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:   
  word_document:
    reference_docx: rv_ms_081921.docx
  pdf_document: default
bibliography: refs.bib
---

@@

# Introduction
Most parasites must navigate a complex environment in order to complete their life cycle and are constrained by the availability and suitability of hosts as well as environmental conditions. Generalist parasites contend with changes in the relative and absolute abundance of their various hosts, and parasites that are transmitted environmentally must rely on favorable conditions for persistence until they are taken up by a new host. Much attention has been paid to anthropogenic-induced changes in the biotic and abiotic conditions experienced by parasites, including introduction and extirpation of species, climate warming, pollution, and land conversion (Altizer et al., 2013; Patz et al., 2004; Young et al., 2017). Many of these changes feed into the dilution effect hypothesis, which posits that transmission (in general, or relative to a focal host species) is augmented as host species richness declines, often in response to anthropogenic influence (Pieter T.J. Johnson et al., 2015; Keesing et al., 2006). However, biotic and abiotic change also occurs naturally over time and space. Moreover, the joint effects of abiotic and biotic drivers of epizootics are rarely considered together (Bienentreu & Lesbarrères, 2020; Dillon & Meentemeyer, 2019; Hall et al., 2010; North et al., 2015; Scott-Baumann & Morgan, 2015), and theoretical prediction for such interaction is underdeveloped. In part, this is due to the notion that transmission is influenced by abiotic factors at large spatial scales and by biotic factors at smaller scales (Halliday & Rohr, 2019; Rohr et al., 2019). However, host mobility often means transmission is occurring at intermediate spatial scales, where communities of hosts are connected, and in these metacommunities, the joint influence of host community structure and environmental conditions on parasite transmission is not well understood. In response to this knowledge gap, we combine theoretical and empirical approaches to study how biotic and abiotic factors interact to promote transmission of ranavirus in amphibian communities. We demonstrate that focusing on the relative abundance of species, in concert with total community size and environmental conditions, provides a more detailed explanation for spatiotemporal variation in outbreak risk that cannot be solely explained by biotic or abiotic variables alone. In addition, we build upon existing theory regarding the relationship between host biodiversity and parasite transmission by demonstrating that host species evenness, rather than richness, can be used to relate community composition to outbreak risk.

>>>>

---
title: "manuscript"
author: "Daniel Suh"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:   
  pdf_document: default
  word_document:
    reference_docx: rv_ms_081921.docx
bibliography: refs.bib
---

@@

# Introduction
Most parasites must navigate a complex environment in order to complete their life cycle and are constrained by the availability and suitability of hosts as well as environmental conditions. Generalist parasites contend with changes in the relative and absolute abundance of their various hosts, and parasites that are transmitted environmentally must rely on favorable conditions for persistence until they are taken up by a new host. Much attention has been paid to anthropogenic-induced changes in the biotic and abiotic conditions experienced by parasites, including introduction and extirpation of species, climate warming, pollution, and land conversion [@altizer_climate_2013; @patz_unhealthy_2004; @young_introduced_2017]. Many of these changes feed into the dilution effect hypothesis, which posits that transmission (in general, or relative to a focal host species) is augmented as host species richness declines, often in response to anthropogenic influence [@johnson_why_2015; @keesing_effects_2006]. However, biotic and abiotic change also occurs naturally over time and space. Moreover, the joint effects of abiotic and biotic drivers of epizootics are rarely considered together [@bienentreu_amphibian_2020; @dillon_direct_2019; @hall_why_2010; @north_anthropogenic_2015; @scott-baumann_review_2015], and theoretical prediction for such interaction is underdeveloped. In part, this is due to the notion that transmission is influenced by abiotic factors at large spatial scales and by biotic factors at smaller scales [@halliday_past_2019; @rohr_towards_2019]. However, host mobility often means transmission is occurring at intermediate spatial scales, where communities of hosts are connected, and in these metacommunities, the joint influence of host community structure and environmental conditions on parasite transmission is not well understood. In response to this knowledge gap, we combine theoretical and empirical approaches to study how biotic and abiotic factors interact to promote transmission of ranavirus in amphibian communities. We demonstrate that focusing on the relative abundance of species, in concert with total community size and environmental conditions, provides a more detailed explanation for spatiotemporal variation in outbreak risk that cannot be solely explained by biotic or abiotic variables alone. In addition, we build upon existing theory regarding the relationship between host biodiversity and parasite transmission by demonstrating that host species evenness, rather than richness, can be used to relate community composition to outbreak risk.
